---
title: |
    | Coursera Finance Specialisation (U Michigan)
    | Course 3 -- Valuation: Alternative Methods
    | Week 1 -- Valuation and Leverage
author: "Joe Nguyen"
date: "15 January 2016"
output: html_document
---

```{r, echo=FALSE}
rm(list = ls())

# Change working directory
dirBase <- "/home/joe/Documents/01-coursera/02-finance"
dirWorking <- "/03-valuation-alt-methods"
setwd(paste0(dirBase, dirWorking))

# Generic finance functions
source(paste0(dirBase, "/01-tvm/r-code/func_finance.R"))
```


## 1.1 Leverage and Cost of Capital

### Cost of Capital: All Equity World

- If a firm/project is **all-equity financed**, the cost of capital (or the return on assets) is the return on equity
- The return on equity is obtained by using the **beta of equity** in **CAPM** for a comparable
- All **published betas** are **betas of equity**


### World has both Equity and Debt

- A firm/project/idea could be financed by equity and debt, and hybrids
- Equity and debt capture the most important aspects of financing for valuation
- **Leverage** means that the firm uses debth in its **capital structure**


### A Path-Breaking Result

- In a world with competitive markets and no frictions, financing (or the mix of equity and debt) has no effects on the value of an idea/project/firm (**Modigliani-Miller**)
    - Value is created by your idea (real asset) and its ability to generate cash flows
    - Financing **cannot** create value
    - The cost of capital, or the return on asset, is determined by the inherent (market) risk (beta) of the business (real asset), not by how it is financed
    

### Debt and Cost of Capital

- Under perfect capital markets, the expected return $\mathbf{E}(r_a)$ is the weighted average of the equity and debt cost of capital, or the weighted average cost of capital (WACC):
$$
WACC = \mathbf{E}(r_a) = \frac{D}{E_L + D} \mathbf{E}(r_d) + \frac{E_L}{E_L + D} \mathbf{E}(r_e^L)
$$
where $\mathbf{E}(r_d)$ is the required rate of return on debt and $\mathbf{E}(r_e^L)$ is the required rate of return on the *levered* equity of the firm.


## 1.2 Leverage and Risk



## 1.4 Valuation: Mega Example Setup

#### An Example

- Online, Inc. is a video gaming company with no debt. Recently, their stock has been added to NASDAQ and it has an (equity) beta of 1.50 (NOTE: this is riskier than market, which has a beta of 1.00)
- Online, Inc. is considering investing in the software business. The business project will require an initial investment of $50 million. If undertaken, the video gaming business will represent 25% of Online, Inc.'s assets
- There is 50% chance the project will generate an annual payoff of $7 million forever, a 40% chance of an annual payoff of $5 million forever, and a 10% chance that the project will fail and generate no cash flows.
- Companies solely in the software business have an equity beta of 1.40. These firms have a debt/equity ratio of 0.25 on average and have riskless debt. Online, Inc. is forecasting that the average market risk premium is about 5% and the risk-free rate on a long-term bond is 4.5%


#### CAPM
$$
r_a = r_f + \beta_a (r_m - r_f)
$$

Here,  
$r_f = 4.5\%$, $r_m = 4.5 + 5 = 9.5\%$, $\beta^{soft}_e = 1.40$, $\beta^{soft}_d = 0$, and $\frac{D}{E} = 0.25$.


## 1.5 Mega Example: Question 1

What is Online, Inc's cost of capital before undertaking the project?

Since Online, Inc. has zero debt, $r_a = r_e$.

```{r}
# Use CAPM
rf <- 0.045
rp <- 0.05
beta <- 1.5
ra <- rf + beta * rp; ra
```


## 1.7 Mega Example: Question 2

What is the IRR of the new project?

SOLUTION:

The expected cash flow accounts for all scenarios (events):  
```{r}
evt.p <- c(0.5, 0.4, 0.1)
evt.c <- c(7e6, 5e6, 0)

c <- evt.p %*% evt.c
cost <- 50e6
IRR(cost, c, inf = TRUE)
```


## 1.9 Mega Example: Question 3

Should Online, Inc. take the new project? (given cost of capital and IRR)

**NO**, since cost of capital (12%) is greater than IRR (11%).

**GUTAM'S ANSWER: Don't know** because IRR is for new *software* project (12%), but cost of capital (r) is for *video gaming* (11%). Hence, we cannot compare these.


## 1.11 Mega Example: Question 4

What is the cost of capital (hurdle rate) of the project (software)?

#### SOLUTION

Asset risk:
$$
\beta_a = \beta_e \left( \frac{E}{D + E} \right) + \beta_d \left( \frac{D}{D + E} \right)
$$
Rearranging for equity risk $\beta_e$:
$$
\beta_e = \beta_a + \frac{D}{E} (\beta_a - \beta_d)
$$
Now for the software industry (superscript $s$),
$$
1.40 = \beta_a^s + 0.25 (\beta_a^s - 0)
$$
So,
$$
\beta_a^s = \frac{1.40}{1.25} = 1.12
$$
Now, use CAPM to get cost of capital of software industry:
$$
r_a^s = r_f + \beta_a^s (r_m - r_f)
$$
```{r}
betaSoft <- 1.40/1.25
rSoft <- rf + betaSoft * rp; rSoft
```


## 1.13 Mega Example: Question 5

Should Online, Inc. take the new project? (given cost of capital of software industry now)

**YES** rSoft < IRR (10.1% < 11%)


## 1.14 Mega Example: Question 6

If Online, Inc. has 1 million shares, what will happen to its stock price?

If Online, Inc. invests in the software product, it will invest $50 million and so reduce earnings per share (EPS). Hence, the stock price will go down.

**Gutam answer: consider NPV**  
For perpetuity,
$$
\begin{align}
\text{NPV} &= I_0 + \text{PV}(c1, c2, ...) \\
           &= I_0 + \frac{c}{r_a^s}
\end{align}
$$

```{r}
# NPV of software project
npv <- -cost + c / rSoft

# price per share (stock price)
pps <- npv / 1e6; pps
```

Share price will **increase** by $4.46.


## 1.16 Mega Example: Question 7

What is Online, Inc.'s new cost of capital? Is this good or bad news?

#### SOLUTION

```{r}
# Previous data - video and software businesses of Online, Inc.
raVideo <- 0.12
betaVideo <- 1.50

raSoft <- 0.101
betaSoft <- 1.12

raOnline <- raVideo * 0.25 + raSoft * 0.75; raOnline
```

While the cost of capital has dropped from 12% to 10.56% (purely video to combination video/software), we cannot say whether or not this is better because we don't have a benchmark to compare to. We can assess the new risk of Online, Inc.

```{r}
betaOnline <- betaVideo * 0.25 + betaSoft * 0.75; betaOnline
```

With a new beta of 1.22 (< 1.50 for video only), we see that lower return comes to lower risk (and vice versa).